Inverse design workflow of implicit surface-based cellular materials via conditional generative model and active learning

### 2025-02-08 manuscript.tex
update on mac Feb 08 2025

\author[1]{Jiaxuan Ma}
\author[3]{Bin Cao}
\author[1]{Yuan Tian\corref{cor1}}
\ead{tian0321@shu.edu.cn}
\author[1,2]{Sheng Sun\corref{cor1}}
\ead{mgissh@t.shu.edu.cn}

@@

%%Research highlights
\begin{highlights}
\item A TPMS-GAN with a navigator was developed to generate TPMS unit cells meeting specific mechanical and geometric conditions.
\item An inversed TPMS-GAN was introduced to design TPMS unit cells directly from geometric features in an explainable way, bypassing the need to interpolate in latent space or adjust implicit function coefficients.
\item A constraint-aware active learning method was proposed to tackle bi-objective optimization problems with imbalanced objective function costs.
\end{highlights}

@@

Adding inequality constraints to BO can be formed by
\begin{equation}
\max_{c(\boldsymbol{x}) \leq \lambda} f(\boldsymbol{x})
\label{eq:11}
\end{equation}
where $c(\boldsymbol{x})$ are the other results of data set $\mathcal{D}$, and it is modeled by another probability model similar to $f(\cdot)$. The $c(\boldsymbol{x})\leq \lambda$ is a constraint function that limited the feasible area in a virtual design space. The constraint-aware EI for a candidate $\hat{\boldsymbol{x}}$ can be written by:
\begin{equation}
I_{\text{EI-C}}(\hat{\boldsymbol{x}})=\Delta_\text{C}(\hat{\boldsymbol{x}})\text{max}\{0, f(\hat{\boldsymbol{x}})-f^*\}=\Delta_\text{C}(\hat{\boldsymbol{x}})I_{\text{EI}}(\hat{\boldsymbol{x}})
    \label{eq:12}
\end{equation}
where $\Delta_\text{C}(\hat{\boldsymbol{x}})\in \{0,1\}$ is a feasibility indicator function that is 1 if $c(\hat{\boldsymbol{x}})\leq \lambda$, and 0 otherwise. 
The distribution of random variable $\Delta_\text{C}(\hat{\boldsymbol{x}})$ is Bernoulli distribution:
\begin{equation}
    PF(\hat{\boldsymbol{x}}):=Pr[c(\boldsymbol{x})\leq \lambda] 
    \label{eq:13}
\end{equation}
The quantity $PF(\hat{\boldsymbol{x}})$ is a cumulative distribution function. These steps lead to the  constraint-aware expected improvement AF:
\begin{equation}
\begin{aligned}
\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}}) &=\mathbb{E}[I_\text{EI-C}(\hat{\boldsymbol{x}})]\\
&= \mathbb{E}[\Delta_{C}(\hat{\boldsymbol{x}})I_{\text{EI}}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\\
&= \mathbb{E}[\Delta_{C}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\mathbb{E}[I_\text{EI}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\\
&=PF(\hat{\boldsymbol{x}})\alpha_{\text{EI}}(\hat{\boldsymbol{x}})
\end{aligned}
\label{eq:14}
\end{equation}
Thus $\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}})$ is the a standard EI of $\hat{\boldsymbol{x}}$ over the feasible point weighted by the probability that $\hat{\boldsymbol{x}}$ is feasible. In the work, we extend the constraint-aware expected improvement AF by surrogate $PF(\hat{\boldsymbol{x}})$ with the normalized distance between the constraint value $c(\hat{\boldsymbol{x}})$ and a target point $c^*$. This distance is mathematically expressed as $ND(||c(\hat{\boldsymbol{x}})-c^*||)$. By incorporating the constraint function, the distribution of \(\alpha_{\text{EI}}(\hat{\boldsymbol{x}})\) is adjusted point-by-point. Notably, this approach eliminates the need to construct an additional probability model to surrogate the constraint function $c(\hat{\boldsymbol{x}}) $. Consequently, Equation. \ref{eq:17} can be reformulated as $\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}}) =ND(||c(\hat{\boldsymbol{x}})-c^*||)\alpha_{\text{EI}}(\hat{\boldsymbol{x}})
$.

@@

We construct a rich database of TMPS unit cell by perturbing the parameters $\{\alpha_1, \alpha_2, t_1, t_2\} $ through Latin hyper-cube sampling (LHS), resulting in 1372 unique structures (exclude some invalid structures). These structures exhibit orthotropic isotropic mechanical properties due to the cubic symmetric property of merged TPMS unit cell. For each structure, the effective mechanical stiffness tensor was computed by numerical homogenization method as illustrated in Section \ref{subsec:homo}. The mechanical properties of the cellular materials are influenced by both the scaffold architecture and the constituent materials. In the work, Ti6Al4V (Ti) was used as the constituent materials for orthopedic implants \citep{Peng2023} (material's parameters are provided in Supplementary Table S1). Ti alloy is bio-inert in human bodies and has been the de facto choice for 3D-printed orthopedic implants, achieving successful clinical applications in repairing bone defects. Figure \ref{fig:2}b shows the mechanical properties' distribution (effective elastic modulus $E$, Possion's ratio $v$) of G, D and Hybird TPMS unit cells (The mechanical properties of G and D were computed for additional 100 samples, which are not included in the dataset of 1372 samples). The distributions of Possion's ratio of G and D exhibit single curve-like distributions with relative density and the Possion's ratio generally decrease as relative density increase, while hybrid TPMS unit cells exhibit significant broader property distribution. The relationship between effective elastic modulus $E$ and relative density $\rho$ follows an exponential form $E \propto E_s\rho^{n}$ (more details are provided in Supplementary Note S1), as founded in previous research \citep{Bauer2017}.

>>>>

\author[1]{Jiaxuan Ma}
\author[3]{Bin Cao}
\author[1]{Yuan Tian\corref{cor1}}
\ead{tian0321@shu.edu.cn}
\author[3]{Tong-Yi Zhang}
\author[1,2]{Sheng Sun\corref{cor1}}
\ead{mgissh@t.shu.edu.cn}

@@

%%Research highlights
\begin{highlights}
\item Developed a TPMS-GAN with a navigator to generate TPMS unit cells meeting specific mechanical and geometric conditions.
\item Developed an inversed TPMS-GAN to design TPMS unit cells directly from geometric features in an explainable way, bypassing the need to interpolate in latent space or adjust implicit function coefficients.
\item Proposed a constraint-aware active learning method to address bi-objective optimization problems with imbalanced objective function costs.
\end{highlights}

@@

Adding inequality constraints to BO can be formed by
\begin{equation}
\max_{c(\boldsymbol{x}) \leq \lambda} f(\boldsymbol{x})
\label{eq:11}
\end{equation}
where $c(\boldsymbol{x})$ are the other results of data set $\mathcal{D}$, and it is modeled by another probability model similar to $f(\cdot)$. The $c(\boldsymbol{x})\leq \lambda$ is a constraint function that limited the feasible area in a virtual design space. The constraint-aware EI for a candidate $\hat{\boldsymbol{x}}$ can be written by:
\begin{equation}
I_{\text{EI-C}}(\hat{\boldsymbol{x}})=\Delta_\text{C}(\hat{\boldsymbol{x}})\text{max}\{0, f(\hat{\boldsymbol{x}})-f^*\}=\Delta_\text{C}(\hat{\boldsymbol{x}})I_{\text{EI}}(\hat{\boldsymbol{x}})
    \label{eq:12}
\end{equation}
where $\Delta_\text{C}(\hat{\boldsymbol{x}})\in \{0,1\}$ is a feasibility indicator function that is 1 if $c(\hat{\boldsymbol{x}})\leq \lambda$, and 0 otherwise. 
The distribution of random variable $\Delta_\text{C}(\hat{\boldsymbol{x}})$ is Bernoulli distribution:
\begin{equation}
    PF(\hat{\boldsymbol{x}}):=Pr[c(\boldsymbol{x})\leq \lambda] 
    \label{eq:13}
\end{equation}
The quantity $PF(\hat{\boldsymbol{x}})$ is a cumulative distribution function. These steps lead to the  constraint-aware expected improvement AF:
\begin{equation}
\begin{aligned}
\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}}) &=\mathbb{E}[I_\text{EI-C}(\hat{\boldsymbol{x}})]\\
&= \mathbb{E}[\Delta_{C}(\hat{\boldsymbol{x}})I_{\text{EI}}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\\
&= \mathbb{E}[\Delta_{C}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\mathbb{E}[I_\text{EI}(\hat{\boldsymbol{x}})|\hat{\boldsymbol{x}}]\\
&=PF(\hat{\boldsymbol{x}})\alpha_{\text{EI}}(\hat{\boldsymbol{x}})
\end{aligned}
\label{eq:14}
\end{equation}
Thus $\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}})$ is the a standard EI of $\hat{\boldsymbol{x}}$ over the feasible point weighted by the probability that $\hat{\boldsymbol{x}}$ is feasible. In the work, we extend the constraint-aware expected improvement AF by replacing $PF(\hat{\boldsymbol{x}})$ with the normalized distance between the constraint value $c(\hat{\boldsymbol{x}})$ and a target point $c^*$. This distance is mathematically expressed as $ND(||c(\hat{\boldsymbol{x}})-c^*||)$. By incorporating the constraint function, the distribution of \(\alpha_{\text{EI}}(\hat{\boldsymbol{x}})\) is adjusted pointwisely. Notably, this approach eliminates the need to construct an additional probability model to surrogate the constraint function $c(\hat{\boldsymbol{x}}) $. Consequently, Equation. \ref{eq:17} can be reformulated as $\alpha_{\text{EI-C}}(\hat{\boldsymbol{x}}) =ND(||c(\hat{\boldsymbol{x}})-c^*||)\alpha_{\text{EI}}(\hat{\boldsymbol{x}})
$.

@@

We construct a rich database of TMPS unit cell by perturbing the parameters $\{\alpha_1, \alpha_2, t_1, t_2\} $ through Latin hyper-cube sampling (LHS), resulting in 1372 unique structures (exclude some invalid structures). These structures exhibit orthotropic isotropic mechanical properties due to the cubic symmetric property of merged TPMS unit cell. For each structure, the effective mechanical stiffness tensor was computed by numerical homogenization method as illustrated in Section \ref{subsec:homo}. The mechanical properties of the cellular materials are influenced by both the scaffold architecture and the constituent materials. In the work, Ti6Al4V (Ti) was used as the constituent materials for orthopedic implants \citep{Peng2023} (material's parameters are provided in Table S1). Ti alloy is bio-inert in human bodies and has been the de facto choice for 3D-printed orthopedic implants, achieving successful clinical applications in repairing bone defects. Figure \ref{fig:2}b shows the mechanical properties' distribution (effective elastic modulus $E$, Possion's ratio $v$) of G, D and Hybird TPMS unit cells (The mechanical properties of G and D were computed for additional 100 samples, which are not included in the dataset of 1372 samples). The distributions of Possion's ratio of G and D exhibit single curve-like distributions with relative density and the Possion's ratio generally decrease as relative density increase, while hybrid TPMS unit cells exhibit significant broader property distribution. The relationship between effective elastic modulus $E$ and relative density $\rho$ follows an exponential form $E \propto E_s\rho^{n}$ (more details are provided in Supplementary Note S1), as founded in previous research \citep{Bauer2017}.
